### 2018-07-24 Optimization-of-String-Rewriting-Operations-for-3D-Fractal-Generation-with-Genetic-Algorithms/presentation.tex
Presentation template was added.

\documentclass{beamer}

\mode<presentation> {
	\usetheme{Berlin}
}

\title[Ninth International Conference on Numerical Methods and Applications, August 20-24, 2018, Borovets, Bulgaria]{
	Optimization of String Rewriting Operations for 3D Fractal Generation with Genetic Algorithms
}

\author{Todor Balabanov, Janeta Sevova, Kolyu Kolev}

\date{20-24.VIII.2018}

\institute[IICT-BAS, NM\&A'18] {
	Institute of Information and Communication Technologies \\ 
	Bulgarian Academy of Sciences \\
	\medskip
	\textit{todorb@iinf.bas.bg}
}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Overview}
\tableofcontents
\end{frame}

\section{Introduction}

\section{Conclusions}

\subsection{Advantages and Disadvantages}

\subsection{Q\&A}

\begin{frame}
\frametitle{Questions and Answers}
\center \huge{Thank you for the attention!}
\end{frame}

\end{document}